Molecular dynamics simulation of synchronization in driven particles

### 2020-12-24 manuscript/manuscript_ajp_2020.tex
cited Taylor, working on model

%-----------------------------------------------------------------
%transition from synchronized to chaotic
%describe types of physical systems - particle types
%%In this work,
%we perform 
%numerical simulations of confined, driven particles
%to model a variety of physical phenomena.
%-----------------------------------------------------------------------
%paper outline
%In the following paper
We describe
our molecular dynamics model in Section~\ref{sec:MD}.
%including various  and confining environments.
We include code to simulate
and visualize the dynamics in this section
and supplementary material.
We summarize
our results 
in Section~\ref{sec:results} 
including synchronized motion of a single confined particle
driven across a periodic landscape in 
Section~\ref{sec:one} and
multiple interacting particles
in Section~\ref{sec:sync},
including stationary propagation of high density kinks
in Sec.~\ref{sec:kink}.
%we
%demonstrate how uniform environments and applied forces
%and create synchronized flow patterns.
We present these results using standard tools of non-linear oscillators
such phase diagrams of velocity versus position.
%PHASE DIAGRAMS SHOWNING COUPLING, OR QUASIPERIODIC COUPLING
%INDICATE COOPERATIVE MOTION IN THE Q1D SUBSTRATE
In Section~\ref{sec:quasiperiod},
we show how an aperiodic landscape modifies the particle dynamics.
Finally we explore the transition to chaotic dynamics in 
Section~\ref{sec:chaos}
and conclude 
in Section~\ref{sec:conclusion}. 
In each section
we suggest exercises for interested students,
as summarized in Section~\ref{sec:problems}.

@@

The simulation is controlled by a $for()$ loop
which runs from an initial to maximum $time$ integer.
Each integer timestep
represents a simulation time of $\Delta t=0.001$.
At each timestep
we evaluate the net force on each particle as a function of its position
$\vec{r}_i(t)$
and then integrate
the equation of motion to move particles
to an updated position
$\vec{r}_i(t+\Delta t)$.
%
The integration is simple
because 
we model the 
particle dynamics 
with an overdamped
equation of motion.
%integrated with the Verlet method.
%The suspending fluid is highly viscous
%and exerts a damping force on the particles equivalent
%to a linear drag force.
%Overdamped dynamics is a common assumption in models of colloidal particles.
%This is an assumption that 
The damping comes from a viscous fluid model
providing a nonconservative force,
modeled as a linear friction $\vec{F}_{drag} = -\eta \vec{v}_i$
sufficient so that
the acceleration of the particle is zero.
Such a model is appropriate when the particles
are small and the viscosity is high [finish... reference...]
This model should be familiar to readers
modeling Milliken's oil drops
in a standard classical mechanics text \cite{Taylor2005}.
%This model is encountered in standard texts,
%for instance Ch. 2 of Taylor's Classical Mechanics.

@@

>>>>

%-----------------------------------------------------------------
%transition from synchronized to chaotic
%describe types of physical systems - particle types
%%In this work,
%we perform 
%numerical simulations of confined, driven particles
%to model a variety of physical phenomena.
%-----------------------------------------------------------------------
%paper outline
%In the following paper
We describe
our molecular dynamics model in Section~\ref{sec:MD}.
%including various  and confining environments.
We include code to simulate
and visualize the landscape and dynamics in this section
and supplementary material.
We summarize
our results 
in Section~\ref{sec:results} 
including synchronized motion of a single confined particle
driven across a periodic landscape in 
Section~\ref{sec:one} and
multiple interacting particles
in Section~\ref{sec:sync},
including stationary propagation of high density kinks
in Sec.~\ref{sec:kink}.
%we
%demonstrate how uniform environments and applied forces
%and create synchronized flow patterns.
We present these results using standard tools of non-linear oscillators
such phase diagrams of velocity versus position.
%PHASE DIAGRAMS SHOWNING COUPLING, OR QUASIPERIODIC COUPLING
%INDICATE COOPERATIVE MOTION IN THE Q1D SUBSTRATE
%In Section~\ref{sec:quasiperiod},
%we show how an aperiodic landscape modifies the particle dynamics.
%Finally we explore the transition to chaotic dynamics in 
%Section~\ref{sec:chaos}
%and conclude 
%in Section~\ref{sec:conclusion}. 
In each section
we suggest exercises for interested students,
as summarized in Section~\ref{sec:problems}.

@@

The simulation is controlled by a $for()$ loop
which runs from an initial to maximum $time$ integer.
Each integer timestep
represents a simulation time of $\Delta t=0.001$.
At each timestep
we evaluate the net force on each particle as a function of its position
$\vec{r}_i(t)$
and then integrate
the equation of motion to move particles
to an updated position
$\vec{r}_i(t+\Delta t)$.
%
The integration is simple
because 
we model the 
particle dynamics 
with an overdamped
equation of motion.
%integrated with the Verlet method.
%The suspending fluid is highly viscous
%and exerts a damping force on the particles equivalent
%to a linear drag force.
%Overdamped dynamics is a common assumption in models of colloidal particles.
%This is an assumption that 
The damping comes from a viscous fluid model
providing a nonconservative force,
modeled as a linear friction $\vec{F}_{drag} = -\eta \vec{v}_i$
sufficient so that
the acceleration of the particle is zero.
Such a model is appropriate when the particles
are small and the viscosity of the suspending medium is high.
% [finish... reference...]
This model should be familiar to readers
modeling Milliken's oil drops
in a standard classical mechanics text \cite{Taylor2005}.
%This model is encountered in standard texts,
%for instance Ch. 2 of Taylor's Classical Mechanics.

@@

      %Molecular dynamics references

    \bibitem{Frenkel2001} D. Frenkel and B. Smit, Understanding Molecular Simulation: From Algorithms to Applications (Academic Press, London, 2001).
    \bibitem{Allen2020} Allen and Tildsley, The Computer Simulation of Liquids.

      \bibitem{Taylor2005} J. Taylor,  Classical mechanics. University Science Books (2005).
